<centre>
<B><font size=6, color="red">ALY6015 - Intermediate Analytics</font>
<BR><font size= 4<I>Team: Predictor </I>
<BR><font size= 4<I>Member: Bhavik, Shreyas and Pravin </I>
<BR>Instructor : Dr. Vladimir Shapiro
<BR><font size =4<I>Project Final Report </I></B></left>


```{r, message=FALSE, warning=FALSE, echo=FALSE}
##Importing the Library
library("plyr")
library("dplyr")
library("tidyr")
library("tibble")
library("plotrix")
library("moments")
library("ggplot2")
library("reshape")
library("lubridate")
library("ISLR")
library("caret")
library("ggplot2")
library("gridExtra")
library("pROC")
library("coefplot")
library("lessR")
library("psych")
library("devtools")
library("car")
library("bestglm")
library("summarytools")
library("kableExtra")
library("glmnet")
library("Metrics")
library("MLmetrics")
library("ggplot2")
library('naniar')
library("car")
library("grid")
library("Matrix")
library("corrplot")
library("MASS")
library(leaps)
```


<B><font size=6, color="#00459a">Introduction</font></B>

The use of data analytics has become increasingly popular in sports. Teams and organizations are utilizing data analysis to improve player performance, game strategy, and overall team success. In this exercise, we will be exploring tennis data. Tennis is a sport played among two or four players, where points are awarded to the player when their opponent fails to return the ball. Using tennis data, we will solve two real-time problems for the tennis community. The first problem involves predicting the duration of a match based on the performance of both players. The match duration will be predicted to assist the broadcast company in determining the number of advertisements that can be shown during the game. We will use linear regression to make the match duration prediction, using player performance as our independent variable. The second problem involves predicting the winner or loser of a match to help the betting company set a basic bet for the player. We will use logistic regression to predict a winner or loser, taking into account factors such as aces, double faults, total games won, and total tiebreakers won.

<B><font size= 5 color="#00459a">Problem Statement 1</font></B>


Predicting match duration based on both player’s match performance data using linear regression method

<B><font size=4.5 >Description :</font></B> 

Reason for solving a problem - XYZ sports channel often broadcasts live tennis matches throughout the year. The advertising that is broadcasted throughout the games is how these stations make money. The marketing team of the channel wanted to attract clients who would like to sell their advertisements. For this, the team wants to know the match time duration beforehand so they might come up with several commercials to air during a game. We would like to help the marketing team by providing them with the approximate match time using prediction model techniques.


<B><font size=4.5 > Method :</font></B> 

Method: Linear Regression

Dependent Variable: Match Duration

Independent Variable: Aces,	Double Faults,	First Serves In,	Break Points Saved,	Break Points Converted,	Total Points Won,	Games Won,	Tiebreaks Won,	Surface

Reason for consideration of Independent Variable is explained below:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Create a data frame with the information
data <- data.frame(
  "Independent Variables" = c("Aces", "Double Faults", "First Serves In", "Break Points Saved", "Break Points Converted", "Total Points Won", "Games Won", "Tiebreaks Won", "Surface"),
  "Definition" = c("Refers to a serve that is not touched by the receiver and results in a point for the server", "A player fails to get their serve into the opposite service box on two consecutive attempts", "First attempt a player makes to serve the ball into the service box", "A situation where the serving player is one point away from winning the game and Opponent saved the point", "Server score a breakpoint", "Number of points scored in a match", "Number of games won in match", "Number of special game won- tiebreak game", "The type of court on which a match is played."),
  "Reason for choosing" = c("More aces less will be the match duration as point is scored in first serve shot", "More DF results in more serves thus higher match duration", "More First serve in results in less over serves thus lower is the match duration", "More Break Point saved would result in more games thus higher match duration", "More Break Point Converted would result in less games thus higher match duration", "More Points would result in high match duration", "More Points would result in high match duration", "More Points would result in high match duration", "Clay surface is easy to play thus more rallies higher match duration")
)

# Create a kable table with the data frame
kable(data, align = c("l", "l", "l"), col =c("Independent Variables","Defination","Impact on Match duration"),
                                             caption = "Information on Independent Variables and Reasons for Choosing them ") %>%
  kable_styling(bootstrap_options = c("striped", "hover","bordered"), full_width = FALSE)
```

<p align = "left"><font size= 1.5> Table 1 - Independent Variable for Problem statement 1 </font></p>
 
<br>

<B><font size= 5 color="#00459a">Problem Statement 2 </font></B> 

Predicting match winner based on players match performance data using logistic regression.

<B><font size=4.5 >Description :</font></B> 

XYZ gambling company often wanted to rightly predict the winner to place a bet. To rightly predict, they wanted to use the analytical approach, thus we will be helping them with a predicted model which will help them to understand the winner beforehand.

<B><font size=4.5 > Method :</font></B> 

Method: Logistic Regression.

Dependent Variable: Winner /Loser

Independent Variables: Aces,	Double Faults,	First Serves In,	Break Points saved,	Break Points Converted,	Total points won,	Game won,	Tiebreaks won

Reason for consideration of Independent Variable is explained below:
```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Create a data frame with the variables and their impact on match
df <- data.frame(
  Variables = c("Aces", "Double Faults", "First Serves In", "Break Points saved", "Break Points Converted", "Total points won", "Game won", "Tiebreaks won"),
  Impact = c("More aces, more probability of winning the match", "More DF, less probability of winning the match", "More first serve in, more probability of winning the match", "Increase in BP saved increases probability of winning the match", "Increase in BP converted increases probability of winning the match", "More Total points won, more probability of winning the match", "More Games won, more probability of winning the match", "More tiebreaks won, more probability of winning the match")
)

# Print the table using kable
kable(df, col.names = c("Variables", "Impact on match"), caption = "Variables and their impact on match")%>%
  kable_styling(bootstrap_options = c("striped", "hover","bordered"), full_width = FALSE, position = "left")

```
<p align = "left"><font size= 1.5> Table 2 - Independent Variable for Problem statement 2 </font></p>

<br>



<B><font size= 5 color="#00459a">Exploratory Data Analysis(EDA)</font></B> 

To address our business query we first need to collect our data. We will be using data from three different sources, so the first step is to merge the datasets. After that, we need to select the relevant variables, and then proceed with data preprocessing, which involves checking and removing null and duplicate values.


Loading the Dataset
```{r Read Dataset, warning=FALSE, message=FALSE, results='hide'}
##Reading Match Details
Match_stats <- read.csv("C:\\Users\\bhavi\\Desktop\\MPS Analytics\\Intermediate Analysis\\Project\\Tennis Data\\Tennis Data\\match_stats_2017_unindexed_csv.csv")

##Reading Tournament Details
Match_detail<- read.csv("C:\\Users\\bhavi\\Desktop\\MPS Analytics\\Intermediate Analysis\\Project\\Tennis Data\\Tennis Data\\match_scores_2017_unindexed_csv.csv")

##Reading Tournament Details
Tournament_detail<-read.csv("C:\\Users\\bhavi\\Desktop\\MPS Analytics\\Intermediate Analysis\\Project\\Tennis Data\\Tennis Data\\tournaments_1877-2017.csv")
head(Tournament_detail)
```

Data Preparation 
```{r Data Preparation, warning=FALSE, message=FALSE, results='hide'}
Match_stats <-merge(Match_stats,Match_detail,by.x = "match_id", by.y = "match_id",
                        all.x = TRUE)
##Keeping only important tournament details
Tournament_detail <- Tournament_detail[, c("tourney_year_id",
                                            "tourney_name",
                                           "tourney_location",
                                           "tourney_month",
                                           "tourney_conditions",
                                           "tourney_surface" )]

##Joining Tournament details to the match data
Match_result <-merge(Match_stats,Tournament_detail,by.x = "tourney_year_id", by.y = "tourney_year_id", all.x = TRUE)

##Dropping irrelevant columns
colnames(Match_result)
Match_result <- Match_result %>% dplyr::select(-c("tourney_order.y",
                                                  "tourney_order.y" , 
                                                  "match_stats_url_suffix.x",
                                                  "match_stats_url_suffix.y" , 
                                                  "match_time", 
                                                  "tourney_location",
                                                  "tourney_slug",
                                                  "tourney_url_suffix" ,
                                                  "tourney_round_name",            
                                                  "match_order",
                                                  "winner_name" ,
                                                  "winner_player_id",
                                                   "winner_slug",
                                                  "loser_name",
                                                  "loser_player_id" ,
                                                  "loser_slug",
                                                  "winner_seed",
                                                  "loser_seed"))

Match_result <- Match_result %>% dplyr::select(-ends_with("total"))
colnames(Match_result)

```


Understanding the data dimension and structure 
```{r Data_Dimension, warning=FALSE, message=TRUE, echo=FALSE}
##Checking the class if each variables in the dataset 
Num_Col <- sum(sapply(Match_result, is.numeric))
Cat_Col <- sum(!sapply(Match_result, is.numeric))
Numer_Obs <- nrow(Match_result)
  
```

```{r warning=FALSE, message=TRUE, echo=FALSE}
##Creating Data Dimension Table
kable(cbind(c("# of Observation", "# of Numeric Variables", "# of Categorical Variables"),
            rbind(Numer_Obs,Num_Col,Cat_Col)), 
      align = "c",
      #col.names = "Count",
      caption = "<font><b><left> Dataset Dimensions</font></b></left> ",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"), full_width = TRUE, position = "left") 
  

```



<p align = "left"><font size= 1.5> Table 3 - Dataset Dimensions </font></p>

There are total 3,815 observations in our dataset and a total of 44 variables out of which 38 are numeric variables and 6 are categorical variables


Checking for NA values
```{r NA Value Check, message=FALSE,warning=FALSE, echo=FALSE, results='hide'}
##Checking the NA values count
kable(sapply(X = Match_result[,colSums(is.na(Match_result)) != 0] , FUN = function(x) sum(is.na(x))), 
      align = "c",
      col.names = "# of NA values",
      caption = "<font><b><left> # of NA values within Variables</font></b></left>",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"))
```


```{r NA plot, message=FALSE,warning=FALSE, echo=FALSE}
##Plot for NA values
# gg_miss_var(Match_result[,colSums(is.na(Match_result)) != 0], show_pct = TRUE) +
# ggtitle("% of NA values ") +
#   geom_text(x = 0.1 , y = 1 , label = "Number of Null Values in each of these variables = 17", 
#             size = 4, hjust = 0.5 , vjust = 0.5)

gg_miss_var(Match_result[,colSums(is.na(Match_result)) != 0], show_pct = TRUE) +
  ggtitle(paste("% of NA values", "\n\nNumber of Null Values in each of these variables = 17")) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1))

```
<p align = "left"><font size= 1.5> Figure 1 - % of NA values within Variables </font></p>

From the table and graph we see that only 13 variables has missing values which is less than 0.5% of the total observation, thus removing the NA rows from the dataset 

```{r Drop NA, message=FALSE,warning=FALSE}
#Dropping NA Values
Match_result <- na.omit(Match_result)
```

Checking for Duplicate rows 
```{r Drop NA 1, message=FALSE,warning=FALSE, echo=FALSE}
##Checking the count for duplicate
kable(cbind("Number of duplicates in data ",sum(duplicated(Match_result$match_id))), 
      align = "c",
      #col.names = "# of NA values",
      caption = "<font><left> # of Duplicate rows</font></left>",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"), position = "left" , full_width = TRUE)
```
<p align = "left"><font size= 1.5> Table 4 - # of Duplicate values in Data set </font></p>

There are 6 duplicates which are less than 0.2% of the the total observation thus, removing duplicates from the data 

```{r Duplicate Check, message=FALSE,warning=FALSE, echo=FALSE}
##Checking the duplicate matches
Checking_Duplicate_match <- Match_result[duplicated(Match_result$match_id),] 

```




```{r Remove Duplicate, message=FALSE,warning=FALSE}
##Removing Duplicates 
Match_result <- Match_result[!duplicated(Match_result$match_id),] 
```
<br>

Now that the data pre-processing has been finished, we can examine how many matches have a recorded match duration, analyze the distribution of match duration, handle any extreme values in match duration as outliers, and then assess the relationship between various variables and match duration. 

In addition, we will examine how the winner and loser of each match perform in terms of match performance metrics, such as aces, double faults, total points won, and so on.

```{r Tournament Detail, warning=FALSE, message=FALSE}
##Counting Unique tournaments and Matches in Data 
kable(cbind(c("# of Tournaments", "# of Matches"),
            rbind(n_distinct(Match_result$tourney_year_id),
      n_distinct(Match_result$match_id))), 
      align = "c",
      #col.names = "Count",
      #caption = "<font><b><center> # of NA values within Variables",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"))

```
<p align = "left"><font size= 1.5> Table 5 - #Tournament and #Matches </font></p> 
The dataset has details of 67 tournaments and ~3,792 matches

Checking the Match Duration Distribution 
```{r,message=FALSE, warning=FALSE}
kable(t(quantile(Match_result$match_duration,probs = seq(0,1,0.05))), 
      align = "c",
      #col.names = "Count", 
      caption = "<font><b><center> Distribution of Match Time ",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"),font_size = 12)

```
<p align = "left"><font size= 1.5> Table 6 - Distribution for Match Duration </font></p> 



```{r,message=FALSE, warning=FALSE}
ggplot(data = Match_result, aes(x = match_duration)) +
  geom_histogram(fill = "tan1", color = "black") +
  labs(x = "Duration(min)", y = "Frequency", title = "Match Duration Distribution") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))



```
<p align = "left"><font size= 1.5>
Figure 2 - Match Duration Distribution</font></p>

From the histogram and quantile distribution, it is observed that most of the matches in our dataset(~55%) has a duration between 60min to 120mins.Only 10% of the matches are played beyond 150mins. There a huge gap between 95% and 100% , thus need to dig deep to understand the outlier in match duration 


```{r Match Duration Details, message=FALSE, warning=FALSE, results='hide'}
##Quantile distribution of match duration
quantile(Match_result$match_duration,probs = seq(0,1,0.05))
quantile(Match_result$match_duration,probs = seq(0.95,1,0.01))

##Calculating upper and lower limit of match duration
Lower_Limit <-  quantile(Match_result$match_duration, 0.25) - 1.5*IQR(Match_result$match_duration)
Upper_Limit <- quantile(Match_result$match_duration, 0.75) + 1.5*IQR(Match_result$match_duration)
```

```{r, message=FALSE, warning=FALSE,echo=FALSE}

kable(t(quantile(Match_result$match_duration,probs = seq(0.95,1,0.01))), 
      align = "c",
      #col.names = "Count", 
      caption = "<font><b><center> Distribution of Match Time ",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"),font_size = 12)
```
<p align = "left"><font size= 1.5> Table 7  -  Distribution of Match Time  for outlier treatment </font></p> 

Looking into quantile(95% to 100%), upper and lower distribution, we see there is huge gap between 98% and 100 thus removing this 2% of the data from the dataset 

```{r, message=FALSE, warning=FALSE}
##Removing Outliers from Match Duration
Match_result <- Match_result[Match_result$match_duration <= 200,]
```

<br>

Given our understanding of the game, we had anticipated the impact of each match performance variable on match duration as stated in the problem statement's method. With the availability of data, we can now examine how match duration is influenced by the independent variable and how match performance metric differ for winners and losers

1- Impact of both players Double Fault on  match duration
```{r, warning=FALSE, message=FALSE, fig.width= 10 }
##Scatter between the Winner Double Faults & Loser Double Faults
plot4<-ggplot(Match_result,aes(winner_double_faults,match_duration,col=Match_result$match_duration))+
  xlab("Double Faults: Winner")+
  ylab("Match Duration")+
  geom_point() +
  labs(color = "Match Duration") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

plot5<-ggplot(Match_result,aes(loser_double_faults,match_duration,col=Match_result$match_duration))+
  xlab("Double Faults: Loser")+
  ylab("Match Duration")+
    geom_point() +
  labs(color = "Match Duration") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))


grid.arrange(plot4,plot5,ncol= 1, 
     top = textGrob("Double Fault vs Match Duration",gp=gpar(fontsize=14,font=1)))

```
<p align = "left"><font size= 1.5>
Figure 3 - Double Faults vs Match Duration</font></p>

From the graph, we can see that there isn't a significant contrast between the number of double faults and the duration of the match for the winning and losing players. However, it's evident that the number of double faults made by both players affects the length of the match. When players commit a low number of double faults, there is a broad range of match durations. On the other hand, when players commit a higher number of double faults, the match duration tends to be longer


2. Impact of Number of Aces on Match duration
```{r, warning=FALSE, message=FALSE, fig.width= 10}
##Scatter Plot grid between the Winner Aces & Loser Aces
plot10<- ggplot(Match_result,aes(Match_result$winner_aces,Match_result$match_duration))+
  xlab("Number of Aces of Winner")+
  ylab("Match Duration")+
  #ggtitle(" Aces of Winner vs Duration of Match") +
  geom_point(col = "tan1") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))



plot11<- ggplot(Match_result,aes(Match_result$loser_aces,Match_result$match_duration))+
  xlab("Number of Aces of Loser")+
  ylab("Match Duration")+
  #ggtitle("Aces of Loser vs Duration of Match")+
  geom_point(col= "lightblue") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))


grid.arrange(plot10,plot11,ncol=2 , 
     top = textGrob("# Aces vs Match Duration",gp=gpar(fontsize=14,font=1)))
```
<p align = "left"><font size= 1.5>
Figure 4 - Aces vs Match Duration</font></p>

From the graph we can see that there no pattern followed by Aces and Match Duration. But it is observed that aces are well distributed for winner from low to high  but for loser the number of aces  are more towards lower number 

3. Impact of Total Points Won on Match Duration 
```{r, warning=FALSE, message=FALSE, fig.width= 10}
##Scatter plot grid between the Winner First Serve Points & Loser First Serve Points
plot6<- ggplot(Match_result,aes(Match_result$winner_total_points_won,Match_result$match_duration))+
  xlab("Total Points Won: Winner")+
  ylab("Match Duration")+
  geom_point(col = "tan1") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))



plot7<- ggplot(Match_result,aes(Match_result$loser_total_points_won,Match_result$match_duration))+
  xlab("Total Points Won:: Loser")+
  ylab("Match Duration")+
   geom_point(col= "lightblue") +
  theme(plot.title = element_text(color = "black", size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))


grid.arrange(plot6,plot7,ncol=2, 
     top = textGrob("Total Points Won vs Match Duration",gp=gpar(fontsize=14,font=1)))

```
<p align = "left"><font size= 1.5>
Figure 5 - Total Point Won vs Match Duration</font></p>

From the graph it is clear that Total Point Won has an impact on match duration. As the Total Point Won increases we see the duration increase. Similar trend is observed for both winner and loser. There is a linear relation between Total Point Won and match duration 

4. Impact of Surface on Match duration
```{r, warning=FALSE, message=FALSE, fig.width= 10 }
##Creating dataset to compare Tournament Surface & Average Time
Surface_Distribution <- Match_result %>% group_by(tourney_surface) %>% summarise( Avg_time = mean(match_duration))
Surface_Distribution$tourney_surface<-as.factor(Surface_Distribution$tourney_surface)

##Plotting barplot
ggplot(Surface_Distribution,aes(tourney_surface,Avg_time,fill=Surface_Distribution$tourney_surface))+
  xlab("Surface of the Tennis Court")+
  ylab("Average Time")+
  geom_bar(stat="identity")+
  ggtitle("Average Time by Surface of Court", ) +
  labs(fill = "Surface") +
  theme(plot.title = element_text(color = "black",hjust = 0.5, size = 14, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))



```
<p align = "left"><font size= 1.5>
Figure 6 - Surface Impact on Match Duration</font></p>
From the graph it is observed that clay surface have slightly higher match duration than other surface 

5. Differences in Losers and Winners Match performance metric
```{r, warning=FALSE, message=FALSE, fig.width= 10, fig.height= 8 }

plot_13 <- data.frame(cbind(c("Winner", "Loser"), rbind(round(mean(Match_result$winner_double_faults),2),  round(mean(Match_result$loser_double_faults),2)))) %>% ggplot(aes(X1,as.numeric(X2))) +
  xlab("")+
  ylab("Average of Double Faults")+
  ggtitle("Average Double Fault Distribution between Winner & Loser")+
  geom_bar(stat="identity",
       fill="tan1")+
  theme(plot.title = element_text(color = "black",hjust = 0.5, size = 11, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))


plot_14 <- data.frame(cbind(c("Winner", "Loser"), rbind(round(mean(Match_result$winner_aces),2),  round(mean(Match_result$loser_aces),2)))) %>% ggplot(aes(X1,as.numeric(X2))) +
  xlab("")+
  ylab("Average of Aces")+
  ggtitle("Average Aces Distribution between Winner & Loser")+
  geom_bar(stat="identity",
       fill="lightblue")+
  theme(plot.title = element_text(color = "black",hjust = 0.5, size = 11, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

plot_15 <- data.frame(cbind(c("Winner", "Loser"), rbind(round(mean(Match_result$winner_total_points_won),2),  round(mean(Match_result$loser_total_points_won),2)))) %>% ggplot(aes(X1,as.numeric(X2))) +
  xlab("")+
  ylab("Average of Total Point Won")+
  ggtitle("Average Total Point Won  Distribution between Winner & Loser")+
  geom_bar(stat="identity",
       fill="tan1")+
  theme(plot.title = element_text(color = "black",hjust = 0.5, size = 11, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

plot_16 <- data.frame(cbind(c("Winner", "Loser"), rbind(round(mean(Match_result$winner_tiebreaks_won),2),  round(mean(Match_result$loser_tiebreaks_won),2)))) %>% ggplot(aes(X1,as.numeric(X2))) +
  xlab("")+
  ylab("Average of Tiebreaks Won")+
  ggtitle("Average Tiebreaks Won Distribution between Winner & Loser")+
  geom_bar(stat="identity",
       fill="lightblue")+
  theme(plot.title = element_text(color = "black",hjust = 0.5, size = 11, face = "bold"),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

grid.arrange(plot_13,plot_14, plot_15,plot_16, ncol=2)
```
<p align = "left"><font size= 1.5>
Figure 7 - Difference in Match Performance metric between winner & loser</font></p>

The graph indicates that winners have a higher number of aces, total points won, and tie break points compared to losers. On the other hand, losers tend to have more double faults. Therefore, it can be inferred that a player's chances of winning are reduced if they have a higher number of double faults, but increased if they score more aces and points.

<br>

In order to make an accurate prediction about the duration of a match, it is crucial to determine the variables that are highly correlated with the match duration. These variables can then be used to develop a model that estimates the expected length of the match. To accomplish this, one possible method is to construct a correlation matrix and examine the correlations between various variables and the match duration. By creating a correlation plot, we can gain insights into how different variables are correlated with the duration of the match.

When predicting the duration of a future match, it may not be known in advance which player will emerge as the winner and which will be the loser. As a result, it is common practice to designate one player as Player 1 and the other as Player 2, regardless of their eventual outcome. This allows for the development of predictive models that do not depend on knowledge of the winner or loser in advance, and can therefore provide accurate estimates of match duration regardless of the eventual outcome.
```{r abbrevating  colunm names, fig.width=25, fig.height= 15} 
##Changing the column names 
colnames(Match_result) <- gsub("^winner", "P1", colnames(Match_result))
colnames(Match_result) <- gsub("^loser", "P2", colnames(Match_result))

##Filtering Numerical Variables 
Numerical_Vraiable <- select_if(Match_result, is.numeric)

```

```{r , fig.width=25, fig.height= 15, message=FALSE, warning=FALSE} 
##Plotting Correlation Matrix
corr <- cor(Numerical_Vraiable)
##Correlation plot
corrplot(corr,  order="original", tl.cex = 1.5)
```
<p align = "left"><font size= 1.5>
Figure 8 - Correlation Matrix</font></p>

From the correlation matrix and plot it was observed that around 21 variables were highly correlated with match duration. The variable with the highest correlation coefficient was the Total Point Won, followed by return game played and service game played. Winner breakpoint converted and round order(match round- final, semi1final, 1st round) share the least correlated coefficient with match duration. Below is the table with top 5 variables  in terms of correlation coefficient with match duration

```{r,message=FALSE, warning=FALSE, echo=FALSE}
##Converting correlation into dataframe
Correlated_variables <- data.frame( corr) %>%
              rownames_to_column() %>%
              gather(key="variable", value="correlation", -rowname)

##Creating a data frame for correlation coefficient between match duration and other variables 
Match_Duration_Correlation <- Correlated_variables[Correlated_variables$rowname == "match_duration" & Correlated_variables$variable != "match_duration",]

##Filtering the variables which share high correlation with match duration
Match_Duration_Correlation <- Match_Duration_Correlation %>% filter(correlation >= 0.7) %>% arrange(desc(correlation))

##Top 5 variables in terms of correlation coefficient with match duration 
Match_Duration_Correlation %>% top_n(10) %>% kable(align = "c",
      #col.names = "Count",
      caption = "<font><b><center> Correlation of Variables with Match Duration",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"))
```
<p align = "left"><font size= 1.5> Table 8  -  Top 5 variables with high correlation coefficient with match duration. </font></p> 


<B><font size= 5 color="#00459a">Building Model for Problem Statement 1 - Linear Regression </font></B> 

To initiate the model building process, we will begin by referring to a correlation matrix. Typically, variables that display a strong correlation with the dependent variable are included in the model. In this case, we have identified 21 variables that have a high correlation with match duration. However, we have observed that many of these variables also have a strong correlation with each other. For instance, "return played" and "serve played" are both strongly correlated with "total points" and "first serve in". Moreover, "set won" is correlated with "game point won". Given our knowledge of the game and the correlation matrices, we have shortlisted a set of independent variables that will be employed in constructing a linear regression model for predicting match duration.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
mat_1 <- c("P1_aces",
          "P1_double_faults",
          "P1_first_serves_in",
          "P1_break_points_saved",
          "P1_break_points_converted",
          "P1_total_points_won",
          "P1_games_won",
          "P1_tiebreaks_won",
          "tourney_conditions",
          "P2_aces",
          "P2_double_faults",
          "P2_first_serves_in",
          "P2_break_points_saved",
          "P2_break_points_converted",
          "P2_total_points_won",
          "P2_games_won",
          "P2_tiebreaks_won",
          "tourney_surface")


# create matrix to hold the data
mat <- matrix(mat_1, ncol = 2, byrow = FALSE)




# create table
kable(mat, align = "c", col.names =  NULL      ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = FALSE, position = "center")


```

<p align = "left"><font size= 1.5> Table 8 - Independent Variables. </font></p> 
The reason for choosing these variables is stated in the Method of Problem Statement 1 


Preparing the dataset based on shortlisted variables
```{r Model Data ,  warning=FALSE, message=FALSE, echo=FALSE}
##Preparing the data for building model
Model_data <- Match_result %>% dplyr::select(c("match_duration",
                                    "P1_aces",
                                    "P1_double_faults",
                                    "P1_first_serves_in",
                                    "P1_break_points_saved",
                                    "P1_break_points_converted",
                                    "P1_total_points_won",
                                    "P2_aces",
                                    "P2_double_faults",
                                    "P2_first_serves_in",
                                    "P2_break_points_saved",
                                    "P2_break_points_converted",
                                    "P2_total_points_won",
                                    "P1_games_won",
                                    "P2_games_won",
                                    "P1_tiebreaks_won",
                                    "P2_tiebreaks_won",
                                    "tourney_surface"
                                    ))

```


It was observed that the playing surface may played a role in determining the duration of a match, as indicated by the plot of Surface vs. Match duration. In order to incorporate this variable into our analysis, we first need to convert the categorical data into numerical form. This can be done using a technique called one-hot encoding, which transforms the categorical variable into a numerical representation that can be used in our analysis.

```{r One Hot Encoding}
##One hot encoding for Surface-Hard
Model_data$Hard_Surface <- case_when(Model_data$tourney_surface == "Hard" ~ 1,TRUE ~ 0)
##One hot encoding for Surface-Hard
Model_data$Clay_Surface<- case_when(Model_data$tourney_surface == "Clay" ~ 1, TRUE ~ 0)

##Predicting Match duration only for Hard and Clay thus dropping matches played on glass
Model_data <- Model_data %>% filter(tourney_surface != "Grass")

##Removing the Categorical variable
Model_data <- Model_data %>% dplyr::select(-c(tourney_surface))
```

We have removed grass as we don't have enough data points to  grass thus proceed the with checking class balance for Surface Conditions with Hard and clay 

```{r, warning =FALSE, message=FALSE, echo=FALSE}
kable(cbind(c("Hard","Clay"),
        rbind(round(((sum(Model_data$Hard_Surface)/3214)*100),2), round(((sum(Model_data$Clay_Surface)/3214)*100),2))),
      align="c",
      caption = "<font><center><b>Class imbalance check for surface(in %) </b></center>",
      digits = 2) %>%
  kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 14)
```

<p align = "left"><font size= 1.5> Table 9 - Class imbalance check for hard & clay surface. </font></p> 

Based on the data in the table, it can be seen that there is an overrepresentation of matches played on hard surfaces, accounting for 65% of the total matches. This class imbalance could skew the results and cause the model to place too much emphasis on hard surfaces, leading to biased predictions. To counteract this issue, it is necessary to address the class imbalance by applying techniques that balance the distribution of classes in the dataset. This will ensure that the model is not overly influenced by any particular class and can make accurate predictions based on the relationships between the predictors and the response variable.

To address the issue we will be using undersampling method where we will reduce the size of the majority class(Hard Surface) to match the size of the minority class(Clay Surface). 

```{r}
##Checking the number of clay surface observations
Clay <- Model_data %>% filter(Clay_Surface == 1)
Hard <- Model_data %>% filter(Hard_Surface == 1)

##Performing Undersampling on Hard surface
Hard_sample  <- sample_n(Hard,nrow(Clay))

##Class balanced data 
Model_data <- rbind(Clay, Hard_sample)

kable(cbind(c("Hard","Clay"),
        rbind(sum(Model_data$Hard_Surface), sum(Model_data$Clay_Surface))),
      align="c",
      caption = "<font><center><b>Class balanced data for surface </b></center>",
      digits = 2) %>%
  kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 14)

Model_data<- Model_data %>% dplyr::select(-c("Hard_Surface"))

```
<p align = "left"><font size= 1.5> Table 10 - Class imbalance data for hard & clay surface. </font></p> 


Splitting the data to build Linear Regression Model 
```{r Split Linear Model}
set.seed(123)
trainIndex <- createDataPartition(y = Model_data$match_duration, p=.70, list = FALSE)

# Loading training data (70% from the original dataset):
train <- Model_data[trainIndex, ]

# Loading test data (30% from the original dataset):
test <- Model_data[-trainIndex, ]

```
The dataset is splitted into 70:30 ration for train and test. The model will be built using train dataset whereas it will be tested for its performance on test data

Training the model with train dataset
```{r Basic Fit of Linear Model}
set.seed(123)
##Building Linear Model 
Fit <- lm(match_duration ~. , data = train )
summary(Fit)
```
The result of a linear model suggest that certain variables had significant impact on the model. These variables include P1_aces, P1_double_faults, P1_total_points_won, P2_aces, P2_double_faults, P2_first_serves_in, P2_break_points_converted, P2_total_points_won, P2_tiebreaks_won, and clay_surface

The model has a high R-squared value of ~93%, indicating a good fit, but it is possible that this high value could be due to multicollinearity, Thus it is important the model for multicollinearity 

Source - https://www.statisticssolutions.com/multicollinearity/ 


Checking multicollinearity
```{r Multicollinearity Check for Basic, fig.width= 4, fig.height= 4}
##Check VIF in the model for multicollinearity 
kable((vif(Fit)),
      align="c",
      caption = "<font><b>Multicollinearity Check</b>",
      digits=2,
      #col.names = "VIF score")
      )%>%
    kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 12, full_width = FALSE) %>%
   row_spec(c(5,6,11,12,13,14), bold = TRUE, italic = TRUE, background =  "#BBBBBB" ) 
 



```
<p align = "left"><font size= 1.5> Table 11 - Multicollinearity Check. </font></p> 


We have used VIF (Variance Inflation Factor) to a measure of multicollinearity between independent variables in a our  model.
A rule of thumb for interpreting the variance inflation factor:

* 1 = not correlated.

* Between 1 and 5 = moderately correlated.

* Greater than 10 = highly correlated.

source - https://www.statisticshowto.com/variance-inflation-factor/

The results of the model indicate that some variables are highly correlated, namely P1_breaks_points_converted, P1_total_points_won, L_break_points_converted, L_total_points_won, P1_games_won, and L_games_won. This high correlation among them suggests that they are causing multicollinearity in the model. Therefore, it is necessary to remove these variables from the model in order to avoid the issue of multicollinearity.

Removing the highly correlated variables
```{r Modify Train and Test for Multi, fig.width=25, fig.height= 15}
train <- train %>% dplyr::select(c("match_duration",
                                    "P1_aces",
                                    "P1_double_faults",
                                    "P1_break_points_saved",
                                    "P1_total_points_won",
                                     "P2_aces",
                                    "P2_double_faults",
                                    "P2_break_points_saved",
                                    "P2_total_points_won",
                                     "P1_tiebreaks_won",
                                    "P2_tiebreaks_won",
                                    "Clay_Surface"))

test <- test %>% dplyr::select(c("match_duration",
                                    "P1_aces",
                                    "P1_double_faults",
                                    "P1_break_points_saved",
                                    "P1_total_points_won",
                                    "P2_aces",
                                    "P2_double_faults",
                                    "P2_break_points_saved",
                                    "P2_total_points_won",
                                    "P1_tiebreaks_won",
                                    "P2_tiebreaks_won",
                                     "Clay_Surface"))

```

In order to get rid of multicollinearity in our model, we have dropped these variables from the model - P1_breaks_points_converted, P2_break_points_converted, P1_games_won,P2_games_won
```{r finalized fitting the model}
set.seed(123)
##Fitting the model after removing the multicollinearity 
Fit <- lm(match_duration ~. , data = train  )
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
set.seed(123)
summary(Fit)

##Check VIF in the model for multicollinearity 
kable((vif(Fit)),
      align="c",
      caption = "<font><b>Multicollinearity Check</b>",
      digits=2,
      col.names = "VIF score") %>%
  kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 14) %>%
   # row_spec(c(5,6,11,12,13,14), bold = TRUE, italic = TRUE, background =  "#BBBBBB" ) %>%
  column_spec(1, width = "30%") %>% column_spec(2, width = "30%")
```
<p align = "left"><font size= 1.5> Table 12 - Multicollinearity Check. </font></p> 

From the summary and VIF table it was observed that after dropping the highly correlated columns, all variables except P1_double_faults, P1_break_points_saved, and P1_tiebreaks_won played a significant role in building the model. Despite the high R-square value of the model, which is 92%, the presence of a strong correlation between P1_points_won and P2_points_won could not be ignored. However, dropping either of these variables would not be an option as it would lead to a loss of information related to the match duration based on the two players involved. However if multicollinearity caused the model to overfit we will address the issue by running a ridge model  after completing the process of feature selection.

Checking RMSE value for Train and Test FIT model
```{r Test Train split}
set.seed(123)
#Loading independent variable for train and test
x_train <- model.matrix(match_duration ~., train)[,-1]
x_test <- model.matrix(match_duration ~., test)[,-1]

#Loading dependent variable for train and test
y_train <- train$match_duration
y_test <- test$match_duration

```

```{r Prediction for basic, warning=FALSE, message=FALSE}
set.seed(123)
##Predicting match duration for train data
pred_train <- predict(Fit,newx = x_train)

##Predicting match duration for train data
pred_test <- predict(Fit,newx = x_test)

##Comparing RMSE value
kable(t(cbind(c("Train","Test"),rbind(rmse(y_train, pred_train),
                                    rmse(y_test, pred_test)))),
      align="c",
      caption="<font><b><center>Train & Test Values</b></center>",
      digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 15)
```
<p align = "left"><font size= 1.5> Table 13 - Match Duration based on Train & Test Data. </font></p> 

The comparison of root mean squared error (RMSE) values between the train and test data revealed a large discrepancy. The RMSE value for the training data was 9 while it was 46 for the test data, indicating that the model was overfitting. To overcome this issue, a method of feature selection is planned to be used in order to remove some columns and prevent overfitting.

```{r, warning=FALSE, message=FALSE, results='hide'}
##Feature Selection through stepwise method
set.seed(123)
Both <- stepAIC(Fit,direction = "both")
Backward <- stepAIC(Fit,direction = "backward")

Both$call
Backward$call
```

The result of the stepwise feature selection method, with a backward direction and both direction was found to have similar results. both techniques suggested dropping P1_tiebreaks_won and P1_breakpoints_saved for a better AIC value



```{r leap, fig.height= 6, fig.width= 10}
##Building best subset model
set.seed(123)
leaps <- regsubsets( match_duration ~ . , data = train)
plot(leaps, scale= "adjr2")

```
<p align = "left"><font size= 1.5>
Figure 9 - Best Subset Model</font></p>

The Best subset selection method suggested to drop P1_tiebreaks_won, P1_breakpoints_saved and P2_breakpoints_saved for better result

```{r model with feature selection, message=FALSE, warning=FALSE, results='hide', echo=FALSE}
##Feature Selection - regsubset
Best_Subset <- lm(match_duration ~ P1_aces + P2_double_faults +  P1_total_points_won + 
    P2_aces + P2_double_faults + P2_total_points_won + P2_break_points_saved  + 
    P2_tiebreaks_won  + Clay_Surface, 
    data = train)

summary(Best_Subset)


```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
##Comparing all AIC values 
kable(cbind(c("Linear Model","Feature Selction Backward", "Feature Selection -Both", "Best_Subset"),rbind(AIC(Fit),AIC(Backward),AIC(Both),AIC(Best_Subset))),
      align = "c",
      caption = "<font><b>Comparison of AIC for Model",
      digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 15)
```
<p align = "left"><font size= 1.5> Table 14 - Comparison of AIC Values </font></p> 


The results from the table suggested that the AIC values were almost the same, so it was expected that the RMSE values would also be similar to the fit model(the initial model). To verify this the backward feature selection model is used to check RMSE for test and train  for overfitting evaluation 


```{r, message=FALSE, warning=FALSE, echo=FALSE}
##Predicting train y through backward selection model and train data
pred_train_bk <- predict(Backward,newx = x_train)

##Predicting train y through backward selection model and train data
pred_test_bk <- predict(Backward,newx = x_test)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
##Comparing RMSE value
kable(t(cbind(c("Train","Test"),rbind(rmse(y_train, pred_train_bk),
                                    rmse(y_test, pred_test_bk)))),
      align="c",
      caption="<font><b><center>Train & Test Values</b></center>",
      digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 15)

```
<p align = "left"><font size= 1.5> Table 15 - Comparison of RMSE Values </font></p> 

The feature selection method produced skewed results as the RMSE values for the train data were around 8.7, whereas for the test data, the RMSE was 46. The large difference between these values indicates overfitting, and therefore, it is necessary to run a ridge model to address this issue.


The result of the analysis showed that there is a strong correlation between the player1 total points won and the player2 total points won in the dataset. This has led to multicollinearity, which has caused the model to overfit. To tackle the issue, ridge regression will be applied. This method adds a penalty term to the regression equation, which reduces the coefficients of correlated variables towards zero and helps in reducing overfitting. In this way, we aim to address the overfitting issue in our case.
```{r Regularization to avoid overfitting}
set.seed(123)
##Reading cv.glmnet function to get optimal lamba value through ridge technique 
ridge_cv <- cv.glmnet(x_train,y_train,alpha = 0)
```

```{r Lambda Coefficent, message=FALSE, warning=FALSE, results='hide'}
##lambda.min value 
ridge_lambdamin <- ridge_cv$lambda.min

##lambda.1se
ridge_lambda1se <- ridge_cv$lambda.1se   

```

```{r, warning=FALSE, message=FALSE, fig.width= 8}
##lamdbamin and Lmabdamax
plot(ridge_cv)
abline(v = log(ridge_lambdamin), col = "turquoise3", lty = 2, lwd = 2)
text(x = log(ridge_lambdamin-0.4), y = 750, label = "lambd.min = 3.17", col = "turquoise3",srt = 90)
abline(v = log(ridge_lambda1se), col = "mediumpurple1", lty = 2, lwd = 2)
text(x = log(ridge_lambda1se+0.5), y = 750, label = "lambd.1se = 4.20", col = "mediumpurple1",srt = 90)
legend("bottomright", legend = c("lambda.min", "lambda.1se"),
       col = c("turquoise3", "mediumpurple1"), lty = c(2, 2), lwd = c(2, 2))
```
<p align = "left"><font size= 1.5>
Figure 10 - Ridge Regression</font></p>

lambdamin of 3.17, it suggests that the regularization parameter value of 3.17 provides gives the minimum cross-validated error among all the lambda values tested. This lambda value represents the model with the best balance between fit and complexity

```{r Rige Model}
set.seed(123)
##Building Ridge Model with lambda,min
model_ridge_min <- glmnet(x_train, y_train, alpha=0, lambda=ridge_cv$lambda.min)
```

```{r}
Ridge_coef <- data.matrix(coef(model_ridge_min))
colnames(Ridge_coef) <- "Coefficient"

Ridge_coef <- t(apply(Ridge_coef, 1, function(x) rev(sort(x))))


kable(t(Ridge_coef),
      align="c",
      caption="<font><b><center>Train & Test Values</b></center>",
      digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 14) %>%
   column_spec(1, width = "5%") %>% column_spec(2, width = "5%")


```
<p align = "left"><font size= 1.5> Table 16 - Ridge Coefficients </font></p> 

From the Ridge model it is observed that coefficient for total point won for both the players has reduced down to 0 due to multicollinearity 

Checking if the overfitting issue is resolved 
```{r}
##Predicting train y through backward selection model and train data
pred_rasso_train <- predict(model_ridge_min,newx = x_train)

##Predicting train y through backward selection model and train data
pred_rasso_test <- predict(model_ridge_min,newx = x_test)
```

```{r,warning=FALSE, message=FALSE}
##Comparing RMSE value
kable(t(cbind(c("Train","Test"),rbind(rmse(y_train, pred_rasso_train),
                                    rmse(y_test, pred_rasso_test)))),
      align="c",
      caption="<font><b><center>Train & Test Values</b></center>",
      digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 15)
```
<p align = "left"><font size= 1.5> Table 17 - Comparison of RMSE Values </font></p> 

The RMSE values for train and test are ~9 which is low thus suggesting a good fit. As both the RMSE values are close our overfitting issue is resolved and we can use ridge model for our prediction 

Creating Ridge model equation 
```{r}
##
##Creating the dataframe to print equation 
Ridge_coef <- t(apply(Ridge_coef, 1, function(x) rev(sort(x))))
var_names <- colnames(Ridge_coef)
equation <- paste(round(Ridge_coef,2), "*", var_names, collapse = " + ")
equation <- paste("Match Duration = ", equation)
 
print(equation)
```
Following equation can be used to predict the match duration. We can use this equation by feeding in players pre-match performance data to predict the future match.The coefficients in the equation shows how match duration changes with one unit change of each independent variable


Our intention is to provide these equations to our stakeholders and ask them to enter the till-date average match performance data for players into the equations,  which will allow them to make predictions about match duration.



<br>

<B><font size= 5 color="#00459a">Building Model for Problem Statement 2 - Logistic Regression </font></B> 

In order to predict the winner and loser, the winner columns(P1) from 50% of the matches are extracted and stored as winner data. These matches were randomly selected through sampling. The rest of the matches are used to determine the loser and the columns corresponding to the losers(P2)are picked and stored as loser data. The losers are assigned a flag of 0, and the winners are assigned a flag of 1. Both the winner and loser data are then combined by appending them together. Finally, this combined dataset is used to understand the factors which will lead to win or lose of the player 

```{r, warning=FALSE, message=FALSE}
set.seed(123)
##Selecting winner data
Winner_data <- sample_n(Match_result, 0.5*nrow(Match_result))

##Loser Data
Loser_data <- subset(Match_result, (Match_result$match_id %in% Winner_data$match_id))


##Filtering only winner columns
Winner_data <- Winner_data %>% dplyr::select(c("match_duration",
                                    "P1_aces",
                                    "P1_double_faults",
                                    "P1_first_serves_in",
                                    "P1_break_points_saved",
                                    "P1_break_points_converted",
                                    "P1_total_points_won",
                                    "P1_games_won",
                                    "P1_tiebreaks_won"))

##Removing winner from column name
colnames(Winner_data) <- gsub("^P1_", "", colnames(Winner_data))

##Filtering only loser columns                            
Loser_data <- Loser_data %>% dplyr::select(c("match_duration",
                                             "P2_aces",
                                   "P2_double_faults",
                                    "P2_first_serves_in",
                                    "P2_break_points_saved",
                                    "P2_break_points_converted",
                                    "P2_total_points_won",
                                    "P2_games_won",
                                    "P2_tiebreaks_won"))
##Removing loser from column name
colnames(Loser_data) <- gsub("^P2_", "", colnames(Loser_data))
                                   
##Assigning winner and loser Winner_flag 
Winner_data$Winner_flag <- 1
Loser_data$Winner_flag <- 0

##Appending winner and Loser data
Logistic_data <- rbind(Winner_data, Loser_data)
Logistic_data <- Logistic_data %>% dplyr::select(-match_duration)

```

To predict winner or loser of the match below mentioned variables will be used as a independent variable

```{r, warning=FALSE, message=FALSE, echo=FALSE}
kable(c("aces",
     "double_faults",
      "first_serves_in",
      "break_points_saved",
      "break_points_converted",
      "total_points_won",
      "games_won",
      "tiebreaks_won"),
       align="c",
      caption = "<font><center><b>Variable for Predicting Winner or Loser Check</b></center>",
      digits=2,
      col.names = "Independent Variable") %>%
  kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 14) %>%
   # row_spec(c(5,6,11,12,13,14), bold = TRUE, italic = TRUE, background =  "#BBBBBB" ) %>%
  column_spec(1, width = "10%")
```
<p align = "left"><font size= 1.5> Table 18 - Independent variable for logistic regression </font></p> 

To avoid the bias in the model lets check the class imbalance in dependent variable
```{r, warning=FALSE, message=FALSE, echo=FALSE}
Logistic_data %>% group_by(Winner_flag) %>% summarise(no_obs = n()) %>%
kable(align="c",
      caption = "<font><center><b>Class imbalance check for win and loss </b></center>",
      digits = 2) %>%
  kable_styling(bootstrap_options = c("basic", "striped", "bordered"), font_size = 14)
```
<p align = "left"><font size= 1.5> Table 19 - Class imbalance check for logistic </font></p> 

From the table it is observed that we have balance data as we have equal observations

```{r, warning=FALSE, message=FALSE}
#Loading independent variable for train and test
set.seed(111)
trainIndex <- createDataPartition(y = Logistic_data$Winner_flag, p=.70, list = FALSE)

train_log <- Logistic_data[trainIndex, ]

# Loading test data (30% from the original dataset):
test_log <- Logistic_data[-trainIndex, ]

#Loading independent variable for train and test
x_train_log <- model.matrix(Winner_flag ~., train_log)[,-1]
x_test_log  <- model.matrix(Winner_flag ~., test_log)[,-1]

#Loading dependent variable for train and test
y_train_log <- train_log$Winner_flag
y_test_log <- test_log$Winner_flag

```

To build a logistic regression the data is splitted into 70% train and 30% test

Building logistic regression using glm function
```{r}
model_1og <- glm(Winner_flag~., data = train_log, family=binomial(link="logit"))
summary(model_1og)
```
In summary, the logistic regression model suggests that all independent variables, with the exception of "aces," are significant in building the model. The model coefficient shows that a higher number of tie_breaks_won, games_won, break_saved, and converted would increase the chances of a player winning, while a higher number of double faults would result in a low probability of a win.  However, the strange finding is that a higher number of first serves in and total points won may lead to low chance of a win.


Using feature selection method on our model to identify the most important features thereby increase the fit of the model and reduce the number of features needed to achieve the desired level of accuracy.

```{r,warning=TRUE,results='hide',message=FALSE}
Both<- step(model_1og ,direction = "both")
```
The feature selection method suggested to drop ace from the model as it provides better AIC, Thus comparing the AIC below

```{r,warning=TRUE,message=FALSE, echo=FALSE}

kable(cbind(c("Logistic Model",
              "Feature Selection -Both"),
            rbind(AIC(model_1og),
                  AIC(Both))),
            align = "c",
            caption = "<font><center><b>Comparison of AIC for Model</center>",
            digits=2) %>%
  kable_styling(bootstrap_options = c("basic","striped","bordered"),font_size = 15)




```
<p align = "left"><font size= 1.5> Table 20 - AIC check for logistic </font></p> 

Based on the information provided in the table, we can conclude that the Feature selection method model has a slightly lower AIC, which suggests a better fit in comparison to the normal model.

Thus we will be removing ace from our model 
```{r, results='hide', warning=FALSE, message=FALSE}
model_1og <- glm(Winner_flag~ double_faults + first_serves_in + 
    break_points_saved + break_points_converted + total_points_won + 
    games_won + tiebreaks_won, data = train_log, family=binomial(link="logit"))
summary(model_1og)
```

Checking the prediction of our model through confusion matrix 
```{r, message=FALSE, warning=FALSE}
##Converting to factor
train_log$Winner_flag <- as.factor(train_log$Winner_flag)
test_log$Winner_flag <- as.factor(test_log$Winner_flag)

##Predicting the win or loss through logistic model and train data 
P_train <- predict(model_1og, newdata = train_log, type = "response")
Ptrain_classes <- as.factor(ifelse(P_train >= 0.5, 1,0))

Confusion_Matrix_train <- confusionMatrix(Ptrain_classes, train_log$Winner_flag)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}

Confusion_Matrix_train$table

```

The results suggest that the model is performing well in terms of accuracy The high number of true positive and true negative values indicates that the model is accurately predicting the outcome for many instances for train data


```{r}
##Predicting the win or loss through logistic model and train data
P_test <- predict(model_1og, newdata = test_log,type = "response")
Ptest_classes <- as.factor(ifelse(P_test >= 0.5, 1,0))

Confusion_Matrix_test <- confusionMatrix(Ptest_classes, test_log$Winner_flag)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}

Confusion_Matrix_test$table

```

The test results too imply that the model is exhibiting good accuracy, as evidenced by the high number of true positive and true negative values. This suggests that the model is able to accurately predict the outcome for a significant number of instances in the text data.


```{r}
Accuracy <- Confusion_Matrix_train$overall[1]
Precision <- Confusion_Matrix_train$byClass[5]
Recall <- Confusion_Matrix_train$byClass[6]
Specificity <- Confusion_Matrix_train$byClass[2]
metrics <- data.frame(
  #Metric = c("Accuracy", "Precision", "Recall", "Specificity"),
  Value = c(Accuracy, Precision, Recall, Specificity)
)

##Creating report for test data
Accuracy_test <- Confusion_Matrix_test$overall[1]
Precision_test <- Confusion_Matrix_test$byClass[5]
Recall_test <- Confusion_Matrix_test$byClass[6]
Specificity_test <- Confusion_Matrix_test$byClass[2]
metrics_test <- data.frame(
  #Metric = c("Accuracy", "Precision", "Recall", "Specificity"),
  Value = c(Accuracy_test, Precision_test, Recall_test, Specificity_test)
)

kable(cbind(c("Train", "Test"),rbind(t(metrics),t(metrics_test))), 
      caption = "Confusion Matrix Result Comparison",
      digits = 2) %>%
   kable_styling(bootstrap_options = c("striped", "hover","bordered"))

```

<p align = "left"><font size= 1.5> Table 21 - Performance Metrics of model </font></p> 

The table displays the performance of a binary classification model on both train and test data, with accuracy, precision, recall, and specificity values reported for both datasets.

The model achieved high accuracy values for both train and test datasets, with the test data accuracy slightly higher than the train data accuracy, suggesting the model is generalizing well. The precision values were also high for both datasets, indicating that the model was good at correctly predicting positive classes. The recall values showed that the model correctly identified a large proportion of positive instances in both datasets.

The specificity values were also high for both datasets, indicating the model was able to correctly identify negative instances.

Overall, the model performed well on both train and test datasets, demonstrating good classification performance.


```{r, warning=FALSE, message=FALSE}
#plotting ROC curve
ROC1 <- roc(test_log$Winner_flag, P_test)
plot(ROC1, print.auc = TRUE, auc.polygon = TRUE, auc.polygon.col = "lightblue", grid = c(0.1, 0.2), xlim = c(1,0), main = "ROC with AUC shaded")
```
<p align = "left"><font size= 1.5>
Figure 11 - ROC Curve</font></p>

The ROC curve is a plot of the true positive rate (TPR) against the false positive rate (FPR). In general, if the curve is close to the top left corner the model has a better fit as it has high specificity associated with high sensitivity. In our model the graph is close to the top left corner thus we can claim our model is better for prediction

AUC is a metric that ranges from 0 to 1, with higher values indicating better performance. An AUC value of 0.988 is close to 1,This suggests that the model is able to differentiate between positive and negative classes with a high degree of accuracy.

Our model shows good performance according to the performance metrics, ROC curve, and AUC, indicating that it can be used for making predictions on new, unseen data. However, before deploying the model, we must validate it using test data to ensure it works well.However, we need to test the model on such data before finalizing it. Currently, the model is fitted with actual outcome results, but we may need to adjust or optimize it after the initial testing.


<B><font size=6, color="#00459a">Conclusion</font></B>

In conclusion, our analysis showed that the number of tiebreaker points won by the both the players and the surface type (Clay) were crucial in determining the match duration. Matches played on Clay tend to last longer, and if the player wins many tiebreaker points, the match is expected to be longer. On the other hand, if both players have high ace points, the match duration is likely to be slower. With an R-square value of 92%, this model can be used for future predictions of match duration. In terms of predicting the winner, except double fault all the other variables were found to be important factors. A player who wins more tiebreakers and has fewer double faults has a higher chance of winning the match. These models can be applied to future matches by inputting the players' per-match historical performance data in the framework.



<B><font size=6, color="#00459a">References</font></B>

* Starner.J. (2017,5th January). Regularization Part 1: Ridge (L2) Regression. https://www.youtube.com/watch?v=Q81RR3yKn30

* Starner.J. (2017,17th January). Regularization Part 2: Lasso (L1) Regression. https://www.youtube.com/watch?v=NGf0voTMlcs

* Starner.J. (2017,29th January). Ridge, Lasso and Elastic-Net Regression in R. https://www.youtube.com/watch?v=ctmNq7FgbvI

* Zach. (2021, 26th August). When to Use Ridge & Lasso Regression. https://www.statology.org/when-to-use-ridge-lasso-regression/

* Lee. S. (2021, 3oth October).lambda.min, lambda.1se and Cross Validation in Lasso  https://www.r-bloggers.com/2021/10/lambda-min-lambda-1se-and-cross-validation-in-lasso-binomial-response/.

* Cho.D. (n.d). Logistic Regression: Feature Selection Methods. https://rpubs.com/ohcsnad/feature_selection_methods 

* Starmer.J. (2022, 23th January). Logistic Regression in R, Clearly Explained!!!!. https://www.youtube.com/watch?v=C4N3_XJJ-jU 

* Yihui. (n.d). The kableExtra package. https://bookdown.org/yihui/rmarkdown-cookbook/kableextra.html